---
title: "The Effectiveness of State Body Camera Policy"
author: "Alex Khater, Rajeev Koneru, and Cora Martin"
date: "`r Sys.Date()`"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

### Introduction

```{r setup, include=FALSE, warning=FALSE, fig.align="center"}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
# knitr::opts_chunk$set(warning = F, results = "hide", message = F)
# knitr::opts_chunk$set(include = F)
# knitr::opts_chunk$set(echo = FALSE)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
always_allow_html = TRUE
```

```{r}
#Calling Packages
library(dplyr, quietly = TRUE)
library(ggplot2)
library(plotly)
library(ggthemes)
library(tidyverse)
library(forecast)
library(ggalluvial)

#Reading the data set
fp1<- data.frame(read.csv("FPS22.csv"))
```

```{r}
#Cleaning null values from data set
fp <- na.omit(fp1)
```

In a 2014 investigation, The Washington Post reported that the FBI had been under-counting fatal police shootings by more than half. An eighteen year-old Michael Brown was killed by police in St. Louis, Missouri, shining a light on the issue of police brutality in the United States. Relying on news, social media, and police reports, the Washington Post compiled a comprehensive list of police shootings from 2015 to 2022 to amend the issue of under-counting police killings of civilians. Police brutality stems from a lack of accountability. The police force is consistently protected from the consequences of violence, regardless of the use of such tools of accountability as police body cameras. We studied the use of body cameras in police shootings to understand whether the use of body cameras was more common in one state over another. More specifically, we asked: within our data set of 6,574 observations of police shootings from 2015 to 2022 in the United States, is there a correlation between the U.S. state of observation and whether a body camera was turned on during the shooting? We found that the use of body cameras during fatal police shootings certainly varies by state, pointing to the conclusion that state policy plays a role in whether body cameras are turned on during active police duty.

This data set includes information on every fatal police shooting from January 1, 2015 to the present. It is updated regularly by the Washington Post. Some of the Post’s key findings from the data set include: 

1. Overall, police nationwide kill approximately 1,000 civilians annually.

2. Black people are killed by police at a higher rate than White people. In absolute numbers, the White population is killed most often by police. Black people, accounting for less than 13 percent of the population, are killed more than twice the rate of White people.

3. The majority of victims are male and between the ages of 20 and 40 years.

4. Police shootings occur in every state. In cities, they are more frequent. The states with the highest rates of shootings (considering state population) are New Mexico, Alaska, and Oklahoma. 

These findings are the result of previous research by the Washington Post. One of the data set’s limitations is  the absence of data on populations at the time of the observation. This information would be valuable for any type of racial injustice analysis. The research we gathered through the EDA began with very basic analyses: 

```{r}
ggplot(fp, aes(x=age)) + geom_histogram( binwidth =0.6,aes(fill = ..count..))+scale_fill_gradient('Count',low='blue',high='red') + labs(x = "Age of Victim", y="Number of Deaths", title = "Age Frequency Histogram")
```

```{r}
ggplot(fp, aes(x=race,fill=race, color=race)) + geom_histogram(stat="count") + labs(x = "Race: Unknown, Asian, Black, Hispanic, Native, Other, White", y="Number of Deaths", title = "Race Frequency Histogram")
```

```{r}
ggplot(fp, aes(x=gender, fill=gender, color=gender)) + geom_histogram(stat="count") + labs(x = "Gender", y="Number of Deaths", title = "Gender Frequency Histogram")
```

```{r}
ggplot(fp, aes(x= manner_of_death, fill=manner_of_death, color=manner_of_death)) + geom_histogram(stat="count") + labs(x = "Manner of Death", y="Number of Deaths", title = "Manner of Death Frequency Histogram")
```

```{r}
ggplot(fp, aes(x=threat_level, fill=threat_level, color=threat_level)) + geom_histogram(stat="count") + labs(x = "Threat Level", y="Number of Deaths", title = "Threat Level Frequency Histogram")
```

These basic univariate analyses were created to better understand the data set and to identify any interesting patterns that emerged. We began doing more complex multivariate analyses and noting any interesting findings:

Hover over the map below to see the breakdown of fatal police shootings, divided by the race of the victim. We looked at the total number of deaths in each state by race and following are some of the insights:

1. We see that the state with the highest level of victims of police violence is California with a total of 885 victims, followed by Texas with a total of 553 and then Florida with 427.

2. These results are consistent with the populations of these states, with the highest being California, then Texas, and then Florida.

3. We also observe that the highest number of deaths is for Hispanic people in California, whereas in Texas and Florida there are more fatal shootings of White people.

```{r}
#Filtering the data by race and states and summarizing it
police_sr <- fp %>%
  filter(!race == "") %>%
  group_by(state, race) %>%
  summarise(deaths = n())
  
#Spreading the data to get deaths by race
spread_sr <- spread(police_sr, race, deaths)

#Getting the total number of deaths per state
spread_sr$Total <- rowSums(spread_sr[,-1], na.rm = TRUE)

#Check the results if needed
# head(spread_sr)

#Data setup for hovering on the map
sr_data <- spread_sr
sr_data$hover <- with(sr_data, paste("Asian", A, '<br>',
                                     "Black",B,'<br>',
                                     "Hispanic",H,'<br>',
                                     "Native American",N,'<br>',
                                     "Other",O,'<br>',
                                     "White",W,'<br>',
                                     "Total Deaths", `Total`))
#Map specifications
graph <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)

#Plotting the graph
plot_ly(spread_sr, z = spread_sr$Total, text = sr_data$hover, locations = sr_data$state, type = 'choropleth',
        locationmode = 'USA-states', color = spread_sr$Total, colors = 'Reds',
        colorbar = list(title = "Counts ")) %>%
        layout(title = 'Number of Victims of Police Violence by Race per State<br>(Hover over the map for a breakdown by race)', geo = graph)
```

Then we looked at the age of the suspect shot, as well as their race. We made the following observations:

1) We see from the boxplot below that the median age for Black people that have been killed by police is 29 years.

2) White people have a relatively higher median age of 35 years whereas Asian people have the highest median age of around 38 years.

```{r}
#Filtering the data by age and race
fp %>%
  filter(!is.na(age) & race != '') %>% #Remove blanks and NAs
  
  #Box plots
  ggplot (aes(x=race, y=age)) +
  geom_boxplot(aes(color = race)) +

  #Axes labels and titles
  labs(x = "Race of Victim", y = "Age of Victim", 
       title = "Distribution of Victims' Age across Race") +
  scale_x_discrete(labels=c('White',
                            'Other',
                            'Native American',
                            'Hispanic',
                            'Black',
                            'Asian')) +
  coord_flip() +
  theme_bw() +
  theme(legend.position = "none")
```

If we look at the age of each victim against the status of their mental health, we can make the following observation: signs of mental illness appear more frequently within the 30s age range while death by police for people age 50 and above are more common for people showing signs of mental illness.

```{r}
#Age against signs of mental illness
ggplot(fp, aes(x=signs_of_mental_illness, y=age, fill=signs_of_mental_illness))+ geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) + labs(x = "Sign of Mental Illness", y="Age", title = "Status of Mental Wellbeing during Fatal Shooting by Age of Victim")
```

We also looked at the death by race and gender, coming up with the following insight: individuals across all races that were shot and killed by police were more often men.

```{r}
#Filtering the data by race and gender
data_by_rg <- fp %>%
                      filter(race != "") %>%
                      filter(gender != "") %>%
                        group_by(race, gender) %>%
                         summarize(No_of_deaths = n())

#Plotting Data
ggplot(data_by_rg, aes(x = race, y = No_of_deaths, fill = gender)) + 
  geom_bar(stat = "identity",position = "dodge") + 
  labs (x = 'Race', y = 'Number of Deaths') +
  ggtitle('Deaths by Race and Gender') +
  scale_x_discrete(labels=c('Asian',
                            'Black',
                            'Hispanic',
                            'Native American',
                            'Other',
                            'White')) +
  theme_few()

```

We then looked at the distribution of deaths by race and the top 5 armed categories. We discovered that around 9% of the Black victims were unarmed whereas only approximately 6% of the White victims were unarmed. Guns were the most used weapon across all races except for Asian individuals. Asian victims were more often wielding knives.

```{r}
# Finding the top 5 arms used by suspects 
top_5_arms <- fp %>%
                group_by(armed) %>%
                summarise(num_arms = n()) %>%
                arrange(desc(num_arms)) %>%
                head(5)

#Filtering the data by top 5 arms found
race_armed_data <- fp %>%
    filter(race != '') %>%
    mutate(armed_mod = ifelse(armed %in% c('gun', 'knife', 'unarmed', 'vehicle', 'undetermined'), as.character(armed), 'Other')) %>%
    group_by(race, armed_mod) %>%
    summarise(Deaths = n())

# Spreading the data 
race_armed_data_spr <- race_armed_data %>%
    spread(armed_mod, Deaths) 

#Replacing missing values with 0
race_armed_data_spr[is.na(race_armed_data_spr)] <- 0 
```

The following graph illustrates the deaths per year by race from 2015-2022:

```{r}
#Plotting the above table in a stacked bar
ggplot(race_armed_data, aes(x = race, y = Deaths, fill = armed_mod)) + 
  geom_bar(stat = "identity",position = "dodge") + 
  labs (x = 'Race', y = 'Number of Deaths') +
  ggtitle('Whether Victims were Armed by Race') +
  scale_x_discrete(labels=c('Asian',
                            'Black',
                            'Hispanic',
                            'Native American',
                            'Other',
                            'White')) + 
   theme_few() 
```

We looked at the distribution of deaths by suspects’ race and whether they were trying to flee or not. The following are some of our most interesting observations:

1) Only 53% of Black victims shot were not fleeing whereas 71% of the Asian victims who were shot were not trying to flee.

2) The car is the most popular method of fleeing among White victims whereas for Black victims, the most popular method of fleeing was by foot. 

```{r}
#Filtering the data and summarizing it  
race_flee_data <- fp %>%
    filter(race != '') %>%
    group_by(race, flee) %>%
    summarise(Deaths = n()) %>%
    spread(flee, Deaths) 

#Replacing the missing values with 0
race_flee_data[is.na(race_flee_data)] <- 0 
```

After performing all of this exploratory data analysis, we decided we wanted to ask a state- and region-based question. Performing a more comparative analysis using this data set could help us understand why police shootings occur in some states over others. This required us to divide the data in a new way. For data visualization purposes, we chose to divide each state into regions. However, our main focus will rely on state-by-state comparative analysis. The following regional subsets include: 

Northwest (NW) | Southwest (SW) | Midwest (MW) | Southeast (SE) | Northeast (NE)
:-: | :-: | :-: | :-: | :-:
California   |   New Mexico   |  Illinois   |  Georgia   |  New York
Washington   |   Arizona   |  Wisconsin   |  Alabama   |  Rhode Island
Oregon   |   Texas   |  Indiana   |  Mississippi   |  Maryland
Nevada   |   Oklahoma   |  Michigan   |  Louisiana   |  Vermont
Idaho   |   Hawaii   |  Minnesota   |  Tennessee   |  Pennsylvania
Utah   |   -   |  Missouri   |  North Carolina   |  Maine
Montana   |   -   |  Iowa   |  South  Carolina   |  New Hampshire
Colorado   |   -   |  Kansas   |  Florida   |  New Jersey
Wyoming   |   -   |  North Dakota   |  Arkansas   |  Connecticut
Arkansas   |   -   |  South Dakota   |  West Virginia   |  Massachusetts
Arkansas   |   -   |  Nebraska   |  DC   |  -
   -     |   -   |  Ohio   |  Virginia   |     -


Fatal shootings in the Northwest United States:

```{r}
#Initial creation of regional subgroups and tests that they cover everything.
NW <- subset(fp, state=="CA"| state=="WA" | state=="OR" | state== "NV"| state=="ID"| state=="UT"|state=="MT"|state=="CO"| state=="WY"| state=="AK")
nrow(NW)
```

Fatal shootings in the Southwest United States:

```{r}
SW <- subset(fp, state== "NM"| state=="AZ"| state== "TX"| state=="OK"| state== "HI")
nrow(SW)
```

Fatal shootings in the Midwest United States:

```{r}
MW <- subset(fp, state== "IL"| state=="WI"| state=="IN"| state=="MI"| state== "MN"| state== "MO"| state== "IA"| state=="KS"| state=="ND"| state== "SD"| state=="NE"| state=="OH")
nrow(MW)
```

Fatal shootings in the Southeast United States:

```{r}
SE <- subset(fp, state == "GA"| state=="AL"| state== "MS"| state=="LA"| state=="TN"| state=="NC"| state=="SC"| state=="FL"| state== "AR"|state=="WV"| state=="DC"| state=="VA"| state=="KY"| state=="DE")
nrow(SE)
```

Fatal shootings in the Northeast United States:

```{r}
NE <- subset(fp, state=="NY"| state== "MD"| state=="RI"| state== "VT"| state=="PA"| state=="ME"| state== "NH"| state=="NJ"| state=="CT"| state=="MA")
nrow(NE)
```

After dividing up the data set, we decided to spotlight the body camera variable. This column in the csv file asked whether the body camera was turned on or off during the fatal police shooting. The data was compiled from official police records, completed by the officers on site. In the Washington Post’s brief article introducing the data set and examining some of the most interesting findings, there was no mention of the body camera variable. This provided us an opportunity to create a relevant, meaningful research project on a little-studied aspect of police work. Our final research question is as follows: 

Within our data set of 6,574 observations of police shootings from 2015 to 2022 in the United States, is there a correlation between the U.S. state of observation and whether a body camera was turned on during the shooting? 

After performing the technical analysis needed to assert that body camera use varies by state, there is still valuable information we need to find a link between state policy on body cameras and the rate at which body cameras are turned on during fatal shootings. The remainder of this paper will attempt to bridge quantitative analysis using the Washington Post data set with qualitative information on state body camera policy. We will also look into funding by state for police work to see if there is a strong correlation between the amount of police funding and the rate at which body cameras are turned on during fatal shootings. 

### Part I: Technical Analysis

```{r}
#Implementation of the subgroups as a new factor called "regions".
fp$regions <- as.factor(ifelse(( fp$state=="CA"| fp$state=="WA" | fp$state=="OR" | fp$state== "NV"| fp$state=="ID"| fp$state=="UT"|fp$state=="MT"|fp$state=="CO"| fp$state=="WY"| fp$state=="AK"), 'NW',
                     ifelse((fp$state== "NM"| fp$state=="AZ"| fp$state== "TX"| fp$state=="OK"| fp$state== "HI"), 'SW', 
                     ifelse((fp$state== "IL"| fp$state=="WI"| fp$state=="IN"| fp$state=="MI"| fp$state== "MN"| fp$state== "MO"| fp$state== "IA"| fp$state=="KS"| fp$state=="ND"| fp$state== "SD"| fp$state=="NE"| fp$state=="OH"), 'MW', 
                     ifelse(( fp$state=="NY"| fp$state== "MD"| fp$state=="RI"| fp$state== "VT"| fp$state=="PA"| fp$state=="ME"| fp$state== "NH"| fp$state=="NJ"| fp$state=="CT"| fp$state=="MA"), 'NE', 'SE')))))
```

```{r}
#Defining 2 new data sets. These show all the variables (with some modification) by state.
sfp <- fp %>%
   group_by(state) %>%
   mutate(stbcp =  sum(body_camera==1)/n())
#This is a numeric proportion of the Men by state
sfp <- sfp %>%
   group_by(state) %>%
   mutate(gen.p =  sum(gender=="M")/n())
#This is a numeric proportion of sings_of_mental_illness by state
sfp <- sfp %>%
   group_by(state) %>%
   mutate(smi.p =  sum(signs_of_mental_illness==TRUE)/n())
#This is a numeric proportion of fleeing victims by state
sfp <- sfp %>%
   group_by(state) %>%
   mutate(flee.p =  sum(flee=="fleeing")/n())
#This is a numeric proportion of attackers by state
sfp <- sfp %>%
   group_by(state) %>%
   mutate(att.p =  sum(threat_level== "attack")/n())
#This is a numeric proportion of armed victims by state
sfp <- sfp %>%
   group_by(state) %>%
   mutate(armed.p =  sum(armed != "unarmed")/n())
#This is a numeric proportion of suspects being shot (as opposed to tasered) by state
sfp <- sfp %>%
   group_by(state) %>%
   mutate(MoD.p =  sum(manner_of_death == "shot")/n())
#This is the average age by state
sfp <- sfp %>%
   group_by(state) %>%
   mutate(age.avg =  sum(age)/n())
#This is variable is for later analysis (maybe for the final) and is the proportion of non-white victims by state
sfp <- sfp %>%
   group_by(state) %>%
   mutate(Non_White_prop =  sum(race != "W")/n())

#This line just drops all non grouped by state data from this version of the data frame
By_State <- subset(sfp, select= -c(name, id, age, latitude, longitude, city, manner_of_death, gender, armed, race, signs_of_mental_illness, flee, threat_level, body_camera, date, is_geocoding_exact))
```

```{r}
rfp <- fp %>%
   group_by(regions) %>%
   mutate(rgbcp =  sum(body_camera==TRUE)/n())

rfp <- sfp %>%
   group_by(regions) %>%
   mutate(gen.p =  sum(gender=="M")/n())

rfp <- sfp %>%
   group_by(regions) %>%
   mutate(smi.p =  sum(signs_of_mental_illness==TRUE)/n())

rfp <- rfp %>%
   group_by(regions) %>%
   mutate(flee.p =  sum(flee=="fleeing")/n())

rfp <- sfp %>%
   group_by(regions) %>%
   mutate(att.p =  sum(threat_level== "attack")/n())
rfp <- sfp %>%
   group_by(regions) %>%
   mutate(armed.p =  sum(armed != "unarmed")/n())
rfp <- sfp %>%
   group_by(regions) %>%
   mutate(MoD.p =  sum(manner_of_death == "shot")/n())
rfp <- sfp %>%
   group_by(regions) %>%
   mutate(age.avg =  sum(age)/n())
rfp <- sfp %>%
   group_by(regions) %>%
   mutate(Non_White_prop =  sum(race != "W")/n())

By_Region <- subset(sfp, select= -c(name, id, age, latitude, longitude, city, manner_of_death, gender, armed, race, signs_of_mental_illness, flee, threat_level, body_camera, regions, date, is_geocoding_exact))
```

Number of fatal shootings where the body camera was on:

```{r}
camera_no <- fp %>% filter(body_camera != TRUE)
camera_yes <- fp %>% filter(body_camera != FALSE)

print(count(camera_yes,body_camera))
```

Number of fatal shootings where the body camera was off:

```{r}
print(count(camera_no,body_camera))
```

We created some figures to visualize body camera use by state and region. In the below bar graph, TRUE signifies a police body camera that was on, while FALSE indicates the body camera was off:

```{r}
#A bar graph showing the number of fatal police shootings AND proportions of body cameras being on by region (too many states to show on a bar graph like this)
ggplot(fp, aes(x = regions, y = frequency(regions), color=body_camera)) + geom_bar(stat = "identity")+ labs(x = "State", y="Number of Deaths", title = "Regional Police Shootings Colored by Body Camera Proportions")
```

A very important variable allowed us to study body camera use variation by state. This variable is stbcp, or the proportion of shootings when a police body camera was turned on by state. We first checked our data for normality. Because the plot below is relatively linear, we concluded this data is close enough to normality for our purposes. 

```{r}
#Normality check of proportions of body cameras being on by state
qqnorm(By_State$stbcp)
#Seems normal
```

This scatter plot below illustrates the comparative power of stbcp. Each point on the graph depicts a state's proportion of shootings where the police body camera was turned on during the incident. These data points also show an interesting regional variation. For example, we can see that there is very little variation in Southwest, and many differences among states in the Midwest.

```{r}
#Scatter Plot of Body Camera Proportion by Region
pl <- ggplot(data=By_State,aes(x = regions,y=stbcp))

pl + geom_point(aes(colour =stbcp,size=4)) + scale_colour_gradient(high='red',low = "blue")
```

We performed a chi-square test to understand whether there is a significant difference between the proportions of body cameras being on during fatal shootings each state. 

\(H_{0}\): There is no significant differences between US States in the proportion of body cameras being turned on during police shootings

\(H_{A}\): There is a significant difference between US State in the proportion of body cameras being turned on during police shootings

Significance Level: \( \alpha = 0.05" \)

```{r}
test<- By_State
#Setting 

cat3 <- test %>% mutate(proportion=cut(stbcp, breaks=c(0, 0.10, 0.2, 0.3, Inf),labels = c("low", "medium", "high", "very high")))

summary(cat3$stbcp)
cat3$state <- as.factor(cat3$state)
cat3$stbcp <- as.factor(cat3$stbcp)

contable= table(cat3$state, cat3$stbcp)
#print(contable)

suppressWarnings({chisq.test(contable)})
```

With a p-value of less 2e-16, we easily pass our significance level of alpha=0.05 and have shown that there exists significant differences between different states' proportions of body camera usage during fatal police shootings. The challenge now becomes, what is causing this variation in state use of body cameras during active police work? We intend to delve into the reasons why there are differences and research what factors may explain these differences between states. This will require understanding state laws and policies regarding the use of police body cameras. We must also understand the police force consequences for turning off body cameras during police activity in different states.

Studying the use of body cameras in police work is an important topic of study for data-driven policy research in the United States. We hope to be able to apply this correlation between the U.S. state of observation and whether the body camera was on or off during the shooting to state policy on body cameras during police work.

### Part II: Results

We have the visualization, as well as the Chi-square test, to confirm that there is a significant difference in stbcp (the proportion of shootings when a police body camera was turned on by state). Let us now look at the rankings by state of stbcp, from most to least likely of having a body camera on during a police shooting.

```{r}
Alabama <- subset(sfp, state=='AL')
#mean(Alabama$stbcp)
Alaska <- subset(sfp, state=='AK')
#mean(Alaska$stbcp)
Arizona <- subset(sfp, state=='AZ')
#mean(Arizona$stbcp)
Arkansas <- subset(sfp, state=='AR')
#mean(Arkansas$stbcp)
California <- subset(sfp, state=='CA')
#mean(California$stbcp)
Colorado <- subset(sfp, state=='CO')
#mean(Colorado$stbcp)
Connecticut <- subset(sfp, state=='CT')
#mean(Connecticut$stbcp)
Delaware <- subset(sfp, state=='DE')
#mean(Delaware$stbcp)
DC <- subset(sfp, state=='DC')
#mean(DC$stbcp)
Florida <- subset(sfp, state=='FL')
#mean(Florida$stbcp)
Hawaii <- subset(sfp, state=='HI')
#mean(Hawaii$stbcp)
Idaho <- subset(sfp, state=='ID')
#mean(Idaho$stbcp)
Illinois <- subset(sfp, state=='IL')
#mean(Illinois$stbcp)
Indiana <- subset(sfp, state=='IN')
#mean(Indiana$stbcp)
Iowa <- subset(sfp, state=='IA')
#mean(Iowa$stbcp)
Kansas <- subset(sfp, state=='KS')
#mean(Kansas$stbcp)
Maine <- subset(sfp, state=='ME')
#mean(Maine$stbcp)
Maryland <- subset(sfp, state=='MD')
#mean(Maryland$stbcp)
Massachusetts <- subset(sfp, state=='MA')
#mean(Massachusetts$stbcp)
Michigan <- subset(sfp, state=='MI')
#mean(Michigan$stbcp)
Minnesota <- subset(sfp, state=='MN')
#mean(Minnesota$stbcp)
Mississippi <- subset(sfp, state=='MI')
#mean(Mississippi$stbcp)
Missouri <- subset(sfp, state=='MO')
#mean(Missouri$stbcp)
Montana <- subset(sfp, state=='MT')
#mean(Montana$stbcp)
Nebraska <- subset(sfp, state=='NE')
#mean(Nebraska$stbcp)
Nevada <- subset(sfp, state=='NV')
#mean(Nevada$stbcp)
New_Hampshire <- subset(sfp, state=='NH')
#mean(New_Hampshire$stbcp)
New_Jersey <- subset(sfp, state=='NJ')
#mean(New_Jersey$stbcp)
New_Mexico <- subset(sfp, state=='NM')
#mean(New_Mexico$stbcp)
New_York <- subset(sfp, state=='NY')
#mean(New_York$stbcp)
North_Carolina <- subset(sfp, state=='NC')
#mean(North_Carolina$stbcp)
North_Dakota <- subset(sfp, state=='ND')
#mean(North_Dakota$stbcp)
Ohio <- subset(sfp, state=='OH')
#mean(Ohio$stbcp)
Oklahoma <- subset(sfp, state=='OK')
#mean(Oklahoma$stbcp)
Oregon <- subset(sfp, state=='OR')
#mean(Oregon$stbcp)
Pennsylvania <- subset(sfp, state=='PA')
#mean(Pennsylvania$stbcp)
Rhode_Island <- subset(sfp, state=='RI')
#mean(Rhode_Island$stbcp)
South_Carolina <- subset(sfp, state=='SC')
#mean(South_Carolina$stbcp)
South_Dakota <- subset(sfp, state=='SD')
#mean(South_Dakota$stbcp)
Tennessee <- subset(sfp, state=='TN')
#mean(Tennessee$stbcp)
Texas <- subset(sfp, state=='TX')
#mean(Texas$stbcp)
Utah <- subset(sfp, state=='UT')
#mean(Utah$stbcp)
Vermont <- subset(sfp, state=='VT')
#mean(Vermont$stbcp)
Virginia <- subset(sfp, state=='VA')
#mean(Virginia$stbcp)
Washington <- subset(sfp, state=='WA')
#mean(Washington$stbcp)
West_Virginia <- subset(sfp, state=='WV')
#mean(West_Virginia$stbcp)
Wisconsin <- subset(sfp, state=='WI')
#mean(Wisconsin$stbcp)
Wyoming <- subset(sfp, state=='WY')
#mean(Wyoming$stbcp)
```

State | stbcp | State | stbcp | State | stbcp
:-: | :-: | :-: | :-: | :-: | :-:
DC   |   0.409   |  Vermont   |  0.333   |  Nevada   |  0.330
Utah   |   0.307   |  Maryland   |  0.277   |  Connecticut   |  0.238
Iowa   |   0.225   |  Hawaii   |  0.212   |  Nebraska   |  0.200
Rhode Island   |   0.200   |  California   |  0.188   |  Idaho   |  0.185
Indiana   |   0.185   |  Oklahoma   |  0.179   |  Arizona   |  0.161
Alabama   |   0.160   |  New York   |  0.160   |  New Mexico   |  0.159
South Carolina   |   0.156   |  New Hampshire   |  0.150   |  Illinois   |  0.148
Wisconsin   |   0.146   |  Virginia   |  0.144   |  North Dakota   |  0.143
North Carolina   |   0.134   |  Michigan   |  0.133   |  Minnesota   |  0.133
Ohio   |   0.131   |  Texas   |  0.131   |  Kansas   |  0.130
Delaware   |   0.125   |  Colorado   |  0.123   |  Louisiana   |  0.119
Washington   |   0.103   |  Kentucky   |  0.103   |  Tennessee   |  0.101
Montana   |   0.100   |  New Jersey   |  0.100   |  Florida   |  0.0995
Georgia   |   0.0988   |  Alaska   |  0.0952   |  Arkansas   |  0.0787
Oregon   |   0.069   |  West Virginia   |  0.0678   |  South Dakota   |  0.0625
Missouri   |   0.037   |  Pennsylvania   |  0.0352   |  Maine   |  0.000
Wyoming   |   0.000

Here we can see that the states with the highest number of observations where the police body camera was turned on include DC, Vermont, and Nevada. The states with the lowest number include Pennsylvania, Maine, and Wyoming--Maine and Wyoming having no recorded police shooting where a body camera was turned on. Let us now do a deep-dive into state policy on body cameras for these six states.

#### District of Columbia



#### Vermont



#### Nevada



#### Pennsylvania



#### Maine



#### Wyoming
