---
title: "Group 7 Final Project"
author: "Alex Khater, Rajeev Koneru, and Cora Martin"
date: "`r Sys.Date()`"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

### Introduction

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
# knitr::opts_chunk$set(warning = F, results = "hide", message = F)
# knitr::opts_chunk$set(include = F)
# knitr::opts_chunk$set(echo = FALSE)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
always_allow_html = TRUE

```

In response to a severe lack of reporting within government sources, The Washington Post compiled a database of every fatal police shooting in the United States from 2015-2022. We are interested in exploring this data, then carrying out research about a specific variable in the data set: the use of body cameras during fatal shootings. 

This exploratory data analysis is divided into five main parts: first, we organize the data; second, we we reshape the data for state- and region-based comparative analyses and build four new variables (police spending, laws that mandate body camera usage, political leaning (Republican or Democrat) in 2020, number of police officers); third, we ask a SMART research question about our data and attempt to answer this question; fourth, we will continue our research by asking a modeling SMART question and attempt to answer this question.

#To skip to the modeling part of this project, please scroll to line 1053, where part 5 starts.


### Part 1: Setting Up the Data

First we call our packages. Then we read the data set that comes from a csv file called FPS22.csv.

```{r}
#Calling Packages
library(dplyr, quietly = TRUE)
library(ggplot2)
library(plotly)
library(ggthemes)
library(tidyverse)
library(forecast)
library(ggalluvial)

#Reading the data set
fp<- data.frame(read.csv("FPS22.csv"))
```

```{r}
#Cleaning null values from data set
fp <- (fp %>% filter(race != ''))
fp <- (fp %>% filter(gender != ''))
camera_no <- fp %>% filter(body_camera != TRUE)
camera_yes <- fp %>% filter(body_camera != FALSE)
fp <- na.omit(fp)
```

After accounting for null values, the data set we are working with has 6,574 observations. Below we have provided a single sample observation:

Name | Date | Manner of Death | Armed | Age | Gender | Race | City
:-: | :-: | :-: | :-: | :-: | :-: | :-: | :-:
 Tim Elliot   |   10/04/2022   |  Shot   |  Gun   |  53   |   M   |  A   |  Shelton


State | Signs of Mental Illness | Threat Level | Flee | Body Camera | Longitude | Latitude | Is Geocoding Exact?
:-: | :-: | :-: | :-: | :-: | :-: | :-: | :-:
 WA   |   1   |  TRUE   |  Not fleeing   |  FALSE   |   -123   |  47.2   |  TRUE

The total number of observations:
```{r}
#Getting row count and latest incident
nrow(fp)
```

```{r}
#Creating new columns from existing columns
fp$date <- as.Date(fp$date, "%m/%d/%Y")
fp$date <- as.Date(fp$date, "%m/%d/%Y")
fp$month <- format(fp$date, format = "%m")
fp$year <- format(fp$date, format = "%Y")
fp$month <- format(fp$date, format = "%m")
fp$year <- format(fp$date, format = "%Y")

```


### Part 3: Reshaping the Data for State and Regional Comparative Analysis

After pursuing the an exploratory data analysis, we decided to do some comparative analyses between states and regions to create a specific, measureable, achievable, relevant, and time-oriented research question to pursue for the remainder of the project.

To do this, we began by dividing the data into regions for easier visualization and comparative analysis. The regions divide each US state as follows:

Northwest (NW) | Southwest (SW) | Midwest (MW) | Southeast (SE) | Northeast (NE)
:-: | :-: | :-: | :-: | :-:
California   |   New Mexico   |  Illinois   |  Georgia   |  New York
Washington   |   Arizona   |  Wisconsin   |  Alabama   |  Rhode Island
Oregon   |   Texas   |  Indiana   |  Mississippi   |  Maryland
Nevada   |   Oklahoma   |  Michigan   |  Louisiana   |  Vermont
Idaho   |   Hawaii   |  Minnesota   |  Tennessee   |  Pennsylvania
Utah   |   -   |  Missouri   |  North Carolina   |  Maine
Montana   |   -   |  Iowa   |  South  Carolina   |  New Hampshire
Colorado   |   -   |  Kansas   |  Florida   |  New Jersey
Wyoming   |   -   |  North Dakota   |  Arkansas   |  Connecticut
Arkansas   |   -   |  South Dakota   |  West Virginia   |  Massachusetts
Arkansas   |   -   |  Nebraska   |  DC   |  -
   -     |   -   |  Ohio   |  Virginia   |     -


Fatal shootings in the Northwest United States:

```{r}
#Initial creation of regional subgroups and tests that they cover everything.
NW <- subset(fp, state=="CA"| state=="WA" | state=="OR" | state== "NV"| state=="ID"| state=="UT"|state=="MT"|state=="CO"| state=="WY"| state=="AK")
nrow(NW)
```

Fatal shootings in the Southwest United States:

```{r}
SW <- subset(fp, state== "NM"| state=="AZ"| state== "TX"| state=="OK"| state== "HI")
nrow(SW)
```

Fatal shootings in the Midwest United States:

```{r}
MW <- subset(fp, state== "IL"| state=="WI"| state=="IN"| state=="MI"| state== "MN"| state== "MO"| state== "IA"| state=="KS"| state=="ND"| state== "SD"| state=="NE"| state=="OH")
nrow(MW)
```

Fatal shootings in the Southeast United States:

```{r}
SE <- subset(fp, state == "GA"| state=="AL"| state== "MS"| state=="LA"| state=="TN"| state=="NC"| state=="SC"| state=="FL"| state== "AR"|state=="WV"| state=="DC"| state=="VA"| state=="KY"| state=="DE")
nrow(SE)
```

Fatal shootings in the Northeast United States:

```{r}
NE <- subset(fp, state=="NY"| state== "MD"| state=="RI"| state== "VT"| state=="PA"| state=="ME"| state== "NH"| state=="NJ"| state=="CT"| state=="MA")
nrow(NE)
```

```{r}
#Implementation of the subgroups as a new factor called "regions".
fp$regions <- as.factor(ifelse(( fp$state=="CA"| fp$state=="WA" | fp$state=="OR" | fp$state== "NV"| fp$state=="ID"| fp$state=="UT"|fp$state=="MT"|fp$state=="CO"| fp$state=="WY"| fp$state=="AK"), 'NW',
                     ifelse((fp$state== "NM"| fp$state=="AZ"| fp$state== "TX"| fp$state=="OK"| fp$state== "HI"), 'SW', 
                     ifelse((fp$state== "IL"| fp$state=="WI"| fp$state=="IN"| fp$state=="MI"| fp$state== "MN"| fp$state== "MO"| fp$state== "IA"| fp$state=="KS"| fp$state=="ND"| fp$state== "SD"| fp$state=="NE"| fp$state=="OH"), 'MW', 
                     ifelse(( fp$state=="NY"| fp$state== "MD"| fp$state=="RI"| fp$state== "VT"| fp$state=="PA"| fp$state=="ME"| fp$state== "NH"| fp$state=="NJ"| fp$state=="CT"| fp$state=="MA"), 'NE', 'SE')))))
```

To determine the likely causality of police body cameras turned on or off, we added several new variables. These will help us understand why there are differences between states in terms of whether body cameras are used during fatal police shootings. We will provide some sample data for each new variable. 

The first new data set we add  is state spending per capita on police for the year 2021.

```{r, results = FALSE}
#This is to make state level data for spending per capita
vspc <- c(477, 1030, 655, 422, 981, 664, 564, 774, 1337, 704, 491, 549, 576, 633, 407, 441, 553, 390, 526, 425, 842, 597, 550, 633, 446, 480, 608, 611, 756, 525, 691, 759, 904, 549, 574, 555, 487, 736, 665, 728, 431, 478, 515, 547,484, 588, 656, 608, 492, 604, 782 )
fp %>% add_column(spendpc = NA)
for (i in 1:nrow(fp)) {
  if(fp$state[i] =="AL"){
    fp$spendpc[i] <- vspc[1]
  }
   if(fp$state[i] =="AK"){
    fp$spendpc[i] <- vspc[2]
   }
   if(fp$state[i] =="AZ"){
    fp$spendpc[i] <- vspc[3]
   }
   if(fp$state[i] =="AR"){
    fp$spendpc[i] <- vspc[4]
   }
   if(fp$state[i] =="CA"){
    fp$spendpc[i] <- vspc[5]
   }
   if(fp$state[i] =="CO"){
    fp$spendpc[i] <- vspc[6]
    }
    if(fp$state[i] =="CT"){
    fp$spendpc[i] <- vspc[7]
    }
    if(fp$state[i] =="DE"){
    fp$spendpc[i] <- vspc[8]
    }
    if(fp$state[i] =="DC"){
    fp$spendpc[i] <- vspc[9]
    }
    if(fp$state[i] =="FL"){
    fp$spendpc[i] <- vspc[10]
    }
    if(fp$state[i] =="GA"){
    fp$spendpc[i] <- vspc[11]
    } 
    if(fp$state[i] =="HI"){
    fp$spendpc[i] <- vspc[12]
    }
    if(fp$state[i] =="ID"){
    fp$spendpc[i] <- vspc[13]
    }
    if(fp$state[i] =="IL"){
    fp$spendpc[i] <- vspc[14]
    }
    if(fp$state[i] =="IN"){
    fp$spendpc[i] <- vspc[15]
    }
    if(fp$state[i] =="IA"){
    fp$spendpc[i] <- vspc[16]
    }
    if(fp$state[i] =="KS"){
    fp$spendpc[i] <- vspc[17]
    }
    if(fp$state[i] =="KY"){
    fp$spendpc[i] <- vspc[18]
    }
    if(fp$state[i] =="LA"){
    fp$spendpc[i] <- vspc[19]
    }
    if(fp$state[i] =="ME"){
    fp$spendpc[i] <- vspc[20]
    }
    if(fp$state[i] =="MD"){
    fp$spendpc[i] <- vspc[21]
    }
    if(fp$state[i] =="MA"){
    fp$spendpc[i] <- vspc[22]
    }
    if(fp$state[i] =="MI"){
    fp$spendpc[i] <- vspc[23]
    }
    if(fp$state[i] =="MN"){
    fp$spendpc[i] <- vspc[24]
    }
    if(fp$state[i] =="MS"){
    fp$spendpc[i] <- vspc[25]
    }
    if(fp$state[i] =="MO"){
    fp$spendpc[i] <- vspc[26]
    }
    if(fp$state[i] =="MT"){
    fp$spendpc[i] <- vspc[27]
    }
    if(fp$state[i] =="NE"){
    fp$spendpc[i] <- vspc[28]
    }
    if(fp$state[i] =="NV"){
    fp$spendpc[i] <- vspc[29]
    }
    if(fp$state[i] =="NH"){
    fp$spendpc[i] <- vspc[30]
    }
    if(fp$state[i] =="NJ"){
    fp$spendpc[i] <- vspc[31]
    }
    if(fp$state[i] =="NM"){
    fp$spendpc[i] <- vspc[32]
    }
    if(fp$state[i] =="NY"){
    fp$spendpc[i] <- vspc[33]
    }
    if(fp$state[i] =="NC"){
    fp$spendpc[i] <- vspc[34]
    }
    if(fp$state[i] =="ND"){
    fp$spendpc[i] <- vspc[35]
    }
    if(fp$state[i] =="OH"){
    fp$spendpc[i] <- vspc[36]
    }
    if(fp$state[i] =="OK"){
    fp$spendpc[i] <- vspc[37]
    }
    if(fp$state[i] =="OR"){
    fp$spendpc[i] <- vspc[38]
    }
    if(fp$state[i] =="PA"){
    fp$spendpc[i] <- vspc[39]
    }
    if(fp$state[i] =="RI"){
    fp$spendpc[i] <- vspc[40]
    }
    if(fp$state[i] =="SC"){
    fp$spendpc[i] <- vspc[41]
    }
    if(fp$state[i] =="SD"){
    fp$spendpc[i] <- vspc[42]
    }
    if(fp$state[i] =="TN"){
    fp$spendpc[i] <- vspc[43]
    }
    if(fp$state[i] =="TX"){
    fp$spendpc[i] <- vspc[44]
    }
    if(fp$state[i] =="UT"){
    fp$spendpc[i] <- vspc[45]
    }
    if(fp$state[i] =="VT"){
    fp$spendpc[i] <- vspc[46]
    }
    if(fp$state[i] =="VA"){
    fp$spendpc[i] <- vspc[47]
    }
    if(fp$state[i] =="WA"){
    fp$spendpc[i] <- vspc[48]
    }
    if(fp$state[i] =="WV"){
    fp$spendpc[i] <- vspc[49]
    }
    if(fp$state[i] =="WI"){
    fp$spendpc[i] <- vspc[50]
    }
    if(fp$state[i] =="WY"){
    fp$spendpc[i] <- vspc[51]
    }
  
  
  
  }

```

Then we add a new binary variable that illustrates whether states have laws that mandate a police officer to use their body camera when interacting with members of the public. Not many states have this particular law, so only Maryland, New Jersey, New Mexico, and South Carolina are given the value 1, while all other states are given the value 0.

```{r, results = FALSE}
#This is to acknowledge the states that (as of 2021) had a laws requiring officers to have the body cameras on during interactions with the public
fp$bclaw <- NA
for (i in 1:nrow(fp)) {
  if(fp$state[i] =="MD"){
    fp$bclaw[i] <- 1
  }
   if(fp$state[i] =="NJ"){
    fp$bclaw[i] <- 1
   }
   if(fp$state[i] =="NM"){
    fp$bclaw[i] <- 1
   }
   if(fp$state[i] =="SC"){
    fp$bclaw[i] <- 1
   }
  else {
    fp$bclaw[i] <- 0
  }
}
```

We add data on the political affiliation of a state. We used the direction the state swung in the 2020 election. That is, either a state voted from President Trump or for President Biden. Negative values indicate a swing towards Trump and positive values indicate a swing towards Biden.

```{r, results = FALSE}
v2020 <- c(-25, -10, 0.31, -28, 29, 14, 87, 20, 19, -3,0.24, 29, -31, 17, -16, -8, -15, -26, -19, 9, 33, 33, 3, 7, -17, -15, -16, -19, 2, 7, 16, 11, 23, -1.3, -33, -8, -33, 16, 1.2, 21, -12, -26, -23,-6, -20, 34, 10, 19, -39, 0.63, -43  )

fp %>% add_column(marg2020 = NA)
for (i in 1:nrow(fp)) {
  if(fp$state[i] =="AL"){
    fp$marg2020[i] <- v2020[1]
  }
   if(fp$state[i] =="AK"){
    fp$marg2020[i] <- v2020[2]
   }
   if(fp$state[i] =="AZ"){
    fp$marg2020[i] <- v2020[3]
   }
   if(fp$state[i] =="AR"){
    fp$marg2020[i] <- v2020[4]
   }
   if(fp$state[i] =="CA"){
    fp$marg2020[i] <- v2020[5]
   }
   if(fp$state[i] =="CO"){
    fp$marg2020[i] <- v2020[6]
    }
    if(fp$state[i] =="CT"){
    fp$marg2020[i] <- v2020[7]
    }
    if(fp$state[i] =="DE"){
    fp$marg2020[i] <- v2020[8]
    }
    if(fp$state[i] =="DC"){
    fp$marg2020[i] <- v2020[9]
    }
    if(fp$state[i] =="FL"){
    fp$marg2020[i] <- v2020[10]
    }
    if(fp$state[i] =="GA"){
    fp$marg2020[i] <- v2020[11]
    } 
    if(fp$state[i] =="HI"){
    fp$marg2020[i] <- v2020[12]
    }
    if(fp$state[i] =="ID"){
    fp$marg2020[i] <- v2020[13]
    }
    if(fp$state[i] =="IL"){
    fp$marg2020[i] <- v2020[14]
    }
    if(fp$state[i] =="IN"){
    fp$marg2020[i] <- v2020[15]
    }
    if(fp$state[i] =="IA"){
    fp$marg2020[i] <- v2020[16]
    }
    if(fp$state[i] =="KS"){
    fp$marg2020[i] <- v2020[17]
    }
    if(fp$state[i] =="KY"){
    fp$marg2020[i] <- v2020[18]
    }
    if(fp$state[i] =="LA"){
    fp$marg2020[i] <- v2020[19]
    }
    if(fp$state[i] =="ME"){
    fp$marg2020[i] <- v2020[20]
    }
    if(fp$state[i] =="MD"){
    fp$marg2020[i] <- v2020[21]
    }
    if(fp$state[i] =="MA"){
    fp$marg2020[i] <- v2020[22]
    }
    if(fp$state[i] =="MI"){
    fp$marg2020[i] <- v2020[23]
    }
    if(fp$state[i] =="MN"){
    fp$marg2020[i] <- v2020[24]
    }
    if(fp$state[i] =="MS"){
    fp$marg2020[i] <- v2020[25]
    }
    if(fp$state[i] =="MO"){
    fp$marg2020[i] <- v2020[26]
    }
    if(fp$state[i] =="MT"){
    fp$marg2020[i] <- v2020[27]
    }
    if(fp$state[i] =="NE"){
    fp$marg2020[i] <- v2020[28]
    }
    if(fp$state[i] =="NV"){
    fp$marg2020[i] <- v2020[29]
    }
    if(fp$state[i] =="NH"){
    fp$marg2020[i] <- v2020[30]
    }
    if(fp$state[i] =="NJ"){
    fp$marg2020[i] <- v2020[31]
    }
    if(fp$state[i] =="NM"){
    fp$marg2020[i] <- v2020[32]
    }
    if(fp$state[i] =="NY"){
    fp$marg2020[i] <- v2020[33]
    }
    if(fp$state[i] =="NC"){
    fp$marg2020[i] <- v2020[34]
    }
    if(fp$state[i] =="ND"){
    fp$marg2020[i] <- v2020[35]
    }
    if(fp$state[i] =="OH"){
    fp$marg2020[i] <- v2020[36]
    }
    if(fp$state[i] =="OK"){
    fp$marg2020[i] <- v2020[37]
    }
    if(fp$state[i] =="OR"){
    fp$marg2020[i] <- v2020[38]
    }
    if(fp$state[i] =="PA"){
    fp$marg2020[i] <- v2020[39]
    }
    if(fp$state[i] =="RI"){
    fp$marg2020[i] <- v2020[40]
    }
    if(fp$state[i] =="SC"){
    fp$marg2020[i] <- v2020[41]
    }
    if(fp$state[i] =="SD"){
    fp$marg2020[i] <- v2020[42]
    }
    if(fp$state[i] =="TN"){
    fp$marg2020[i] <- v2020[43]
    }
    if(fp$state[i] =="TX"){
    fp$marg2020[i] <- v2020[44]
    }
    if(fp$state[i] =="UT"){
    fp$marg2020[i] <- v2020[45]
    }
    if(fp$state[i] =="VT"){
    fp$marg2020[i] <- v2020[46]
    }
    if(fp$state[i] =="VA"){
    fp$marg2020[i] <- v2020[47]
    }
    if(fp$state[i] =="WA"){
    fp$marg2020[i] <- v2020[48]
    }
    if(fp$state[i] =="WV"){
    fp$marg2020[i] <- v2020[49]
    }
    if(fp$state[i] =="WI"){
    fp$marg2020[i] <- v2020[50]
    }
    if(fp$state[i] =="WY"){
    fp$marg2020[i] <- v2020[51]
    }
  
  
  
  }

```

Finally, we add a variable that looks at a state's quantity of police officers per 100K citizens.

```{r, results = FALSE}
vle <- c(418.356, 604.3172, 507.1749, 455.5134, 378.4166, 416.7751, 406.803, 517.3799, 722, 439.7941,  487.2325,494.1445, 378.7875, 476.7941, 379.4256, 307.9675, 466.5578, 390.2951, 595.0071, 411.2744,619.3873,440.4483,359.6753,324.5891,487.8593,459.6233,435.4656, 412.0619, 409.9609, 419.7185, 532.2092, 506.5776, 630.0293, 438.4865, 392.6816, 395.1726, 410.4272, 283.7333, 462.605, 412.0791, 420.8136, 449.0854, 438.823, 479.3943, 299.0369, 321.71, 597.2275, 319.9674, 516.3705, 404.1405, 501.126)

fp %>% add_column(le_per_100k = NA)
for (i in 1:nrow(fp)) {
  if(fp$state[i] =="AL"){
    fp$le_per_100k[i] <- vle[1]
  }
   if(fp$state[i] =="AK"){
    fp$le_per_100k[i] <- vle[2]
   }
   if(fp$state[i] =="AZ"){
    fp$le_per_100k[i] <- vle[3]
   }
   if(fp$state[i] =="AR"){
    fp$le_per_100k[i] <- vle[4]
   }
   if(fp$state[i] =="CA"){
    fp$le_per_100k[i] <- vle[5]
   }
   if(fp$state[i] =="CO"){
    fp$le_per_100k[i] <- vle[6]
    }
    if(fp$state[i] =="CT"){
    fp$le_per_100k[i] <- vle[7]
    }
    if(fp$state[i] =="DE"){
    fp$le_per_100k[i] <- vle[8]
    }
    if(fp$state[i] =="DC"){
    fp$le_per_100k[i] <- vle[9]
    }
    if(fp$state[i] =="FL"){
    fp$le_per_100k[i] <- vle[10]
    }
    if(fp$state[i] =="GA"){
    fp$le_per_100k[i] <- vle[11]
    } 
    if(fp$state[i] =="HI"){
    fp$le_per_100k[i] <- vle[12]
    }
    if(fp$state[i] =="ID"){
    fp$le_per_100k[i] <- vle[13]
    }
    if(fp$state[i] =="IL"){
    fp$le_per_100k[i] <- vle[14]
    }
    if(fp$state[i] =="IN"){
    fp$le_per_100k[i] <- vle[15]
    }
    if(fp$state[i] =="IA"){
    fp$le_per_100k[i] <- vle[16]
    }
    if(fp$state[i] =="KS"){
    fp$le_per_100k[i] <- vle[17]
    }
    if(fp$state[i] =="KY"){
    fp$le_per_100k[i] <- vle[18]
    }
    if(fp$state[i] =="LA"){
    fp$le_per_100k[i] <- vle[19]
    }
    if(fp$state[i] =="ME"){
    fp$le_per_100k[i] <- vle[20]
    }
    if(fp$state[i] =="MD"){
    fp$le_per_100k[i] <- vle[21]
    }
    if(fp$state[i] =="MA"){
    fp$le_per_100k[i] <- vle[22]
    }
    if(fp$state[i] =="MI"){
    fp$le_per_100k[i] <- vle[23]
    }
    if(fp$state[i] =="MN"){
    fp$le_per_100k[i] <- vle[24]
    }
    if(fp$state[i] =="MS"){
    fp$le_per_100k[i] <- vle[25]
    }
    if(fp$state[i] =="MO"){
    fp$le_per_100k[i] <- vle[26]
    }
    if(fp$state[i] =="MT"){
    fp$le_per_100k[i] <- vle[27]
    }
    if(fp$state[i] =="NE"){
    fp$le_per_100k[i] <- vle[28]
    }
    if(fp$state[i] =="NV"){
    fp$le_per_100k[i] <- vle[29]
    }
    if(fp$state[i] =="NH"){
    fp$le_per_100k[i] <- vle[30]
    }
    if(fp$state[i] =="NJ"){
    fp$le_per_100k[i] <- vle[31]
    }
    if(fp$state[i] =="NM"){
    fp$le_per_100k[i] <- vle[32]
    }
    if(fp$state[i] =="NY"){
    fp$le_per_100k[i] <- vle[33]
    }
    if(fp$state[i] =="NC"){
    fp$le_per_100k[i] <- vle[34]
    }
    if(fp$state[i] =="ND"){
    fp$le_per_100k[i] <- vle[35]
    }
    if(fp$state[i] =="OH"){
    fp$le_per_100k[i] <- vle[36]
    }
    if(fp$state[i] =="OK"){
    fp$le_per_100k[i] <- vle[37]
    }
    if(fp$state[i] =="OR"){
    fp$le_per_100k[i] <- vle[38]
    }
    if(fp$state[i] =="PA"){
    fp$le_per_100k[i] <- vle[39]
    }
    if(fp$state[i] =="RI"){
    fp$le_per_100k[i] <- vle[40]
    }
    if(fp$state[i] =="SC"){
    fp$le_per_100k[i] <- vle[41]
    }
    if(fp$state[i] =="SD"){
    fp$le_per_100k[i] <- vle[42]
    }
    if(fp$state[i] =="TN"){
    fp$le_per_100k[i] <- vle[43]
    }
    if(fp$state[i] =="TX"){
    fp$le_per_100k[i] <- vle[44]
    }
    if(fp$state[i] =="UT"){
    fp$le_per_100k[i] <- vle[45]
    }
    if(fp$state[i] =="VT"){
    fp$le_per_100k[i] <- vle[46]
    }
    if(fp$state[i] =="VA"){
    fp$le_per_100k[i] <- vle[47]
    }
    if(fp$state[i] =="WA"){
    fp$le_per_100k[i] <- vle[48]
    }
    if(fp$state[i] =="WV"){
    fp$le_per_100k[i] <- vle[49]
    }
    if(fp$state[i] =="WI"){
    fp$le_per_100k[i] <- vle[50]
    }
    if(fp$state[i] =="WY"){
    fp$le_per_100k[i] <- vle[51]
    }
  
  
  
  }
```

Example data points are shown below:

State | Variable | Value
:-: | :-: | :-:
Alabama   |   Police Spending Per Capita   |  USD$477
Maryland   |   Police Body Camera Laws   |  1
Iowa   |   2020 US Presidential Election Vote   |  -8
Florida   |   Police Quantity Per 100k Citizens   |  USD$477

We also created two sub-data sets by grouping the data by state and by region for visualization purposes. The contents of both groups are identical, besides their grouping.

```{r, results=FALSE}
ffp <- subset(fp, (year== 2022| year==2021) )
```

```{r}
#Defining 2 new data sets. These show all the variables (with some modification) by state.

sfp <- ffp %>%
   group_by(state) %>%
   mutate(stbcp =  sum(body_camera==1)/n())
#This is a numeric proportion of the Men by state
sfp <- sfp %>%
   group_by(state) %>%
   mutate(gen.p =  sum(gender=="M")/n())
#This is a numeric proportion of sings_of_mental_illness by state
sfp <- sfp %>%
   group_by(state) %>%
   mutate(smi.p =  sum(signs_of_mental_illness==TRUE)/n())
#This is a numeric proportion of fleeing victims by state
sfp <- sfp %>%
   group_by(state) %>%
   mutate(flee.p =  sum(flee=="fleeing")/n())
#This is a numeric proportion of attackers by state
sfp <- sfp %>%
   group_by(state) %>%
   mutate(att.p =  sum(threat_level== "attack")/n())
#This is a numeric proportion of armed victims by state
sfp <- sfp %>%
   group_by(state) %>%
   mutate(armed.p =  sum(armed != "unarmed")/n())
#This is a numeric proportion of suspects being shot (as opposed to tasered) by state
sfp <- sfp %>%
   group_by(state) %>%
   mutate(MoD.p =  sum(manner_of_death == "shot")/n())
#This is the average age by state
sfp <- sfp %>%
   group_by(state) %>%
   mutate(age.avg =  sum(age)/n())
#This is variable is for later analysis (maybe for the final) and is the proportion of non-white victims by state
sfp <- sfp %>%
   group_by(state) %>%
   mutate(Non_White_prop =  sum(race != "W")/n())

#This line just drops all non grouped by state data from this version of the data frame
By_State <- subset(sfp, select= -c(name, id, age, latitude, longitude, city, manner_of_death, gender, armed, race, signs_of_mental_illness, flee, threat_level, body_camera, date, is_geocoding_exact))
```

```{r}
rfp <- sfp %>%
   group_by(regions) %>%
   mutate(rgbcp =  sum(body_camera==TRUE)/n())

rfp <- sfp %>%
   group_by(regions) %>%
   mutate(gen.p =  sum(gender=="M")/n())

rfp <- sfp %>%
   group_by(regions) %>%
   mutate(smi.p =  sum(signs_of_mental_illness==TRUE)/n())

rfp <- rfp %>%
   group_by(regions) %>%
   mutate(flee.p =  sum(flee=="fleeing")/n())

rfp <- sfp %>%
   group_by(regions) %>%
   mutate(att.p =  sum(threat_level== "attack")/n())
rfp <- sfp %>%
   group_by(regions) %>%
   mutate(armed.p =  sum(armed != "unarmed")/n())
rfp <- sfp %>%
   group_by(regions) %>%
   mutate(MoD.p =  sum(manner_of_death == "shot")/n())
rfp <- sfp %>%
   group_by(regions) %>%
   mutate(age.avg =  sum(age)/n())
rfp <- sfp %>%
   group_by(regions) %>%
   mutate(Non_White_prop =  sum(race != "W")/n())

By_Region <- subset(sfp, select= -c(name, id, age, latitude, longitude, city, manner_of_death, gender, armed, race, signs_of_mental_illness, flee, threat_level, body_camera, regions, date, is_geocoding_exact))
```

### Part 4: Research SMART Question and Answer

Within our data set of 5,720 observations of police shootings from 2015 to 2022 in the United States, is there a correlation between the U.S. state of observation and whether a body camera was turned on during the shooting?

The state data subgroup can be summarized as follows:

```{r}
#Just some tests on the new sub data set
summary(By_State)
```

The region data subgroup can be summarized as follows:

```{r}
summary(By_Region)
```

#### Figure 1: Normality

We will now check our data for normality:

```{r}
#Normality check of proportions of body cameras being on by state
qqnorm(By_State$stbcp)
#Seems normal
```

Because the plot is relatively linear, we can conclude this data is close enough to normality for our purpose. 

#### Figure 2: Body Camera Usage by Region

Now let us look at the body camera proportions by state. In the below bar graph, TRUE signifies a police body camera that was on, while FALSE indicates the body camera was off:

```{r}
#A bar graph showing the number of fatal police shootings AND proportions of body cameras being on by region (too many states to show on a bar graph like this)
ggplot(fp, aes(x = regions, y = frequency(regions), color=body_camera)) + geom_bar(stat = "identity")+ labs(x = "State", y="Number of Deaths")
```

Number of fatal shootings where the body camera was on:

```{r}

print(count(camera_yes,body_camera))
```

Number of fatal shootings where the body camera was off:

```{r}
print(count(camera_no,body_camera))
```

#### Figure 3: stbcp by Region

This scatter plot shows the proportion of fatal shootings when cameras were on by state (the variable stbcp). Each point on the graph depicts a state's proportion of shootings where the police body camera was turned on during the incident). We can see that there is very little variation in Southwest, and many differences among states in the Midwest.

```{r}
#Scatter Plot of Body Camera Proportion by Region
pl <- ggplot(data=By_State,aes(x = regions,y=stbcp))

pl + geom_point(aes(colour =stbcp,size=4)) + scale_colour_gradient(high='red',low = "blue")
```

Finally, let us check out the mean body camera on proportion for all states:

```{r}
mean(By_State$stbcp)
```

And the stbcp median body camera on proportion for all states:

```{r}
median(By_State$stbcp)
```

#### Model 1: Chi-square Test

We will now perform a chi-square test to see if there is a significant difference between the proportions of each state. 

Null: There is no significant differences between US States in the proportion of body cameras being turned on during police shootings

Alternative: There is a significant difference between US State in the proportion of body cameras being turned on during police shootings

Significance Level: a = 0.05

```{r}
test<- By_State
#Setting 

cat3 <- test %>% mutate(proportion=cut(stbcp, breaks=c(0, 0.10, 0.2, 0.3, Inf),labels = c("low", "medium", "high", "very high")))

summary(cat3$stbcp)
cat3$state <- as.factor(cat3$state)
cat3$stbcp <- as.factor(cat3$stbcp)

contable= table(cat3$state, cat3$stbcp)
#print(contable)

suppressWarnings({chisq.test(contable)})
```

With a p-value of 2e-16, we easily pass our significance level of alpha=0.05 and have shown that there exists significant differences between different states' proportions of body camera usage during fatal police shootings.

### Part 5: Investigating Causality

We now know there are significant differences in the level of body camera usage during police shootings among US states. Let us see if we can find out what drives those differences. 

Our second SMART question: For the years 2021 and 2022, what variables influence a state's proportion of body cameras turned on during fatal police shootings? 

The variables we will study include: 

1. US region

2. Law enforcement officers per 100,000 citizens

3. Law enforcement spending per capita

4. Body camera mandate laws

5. 2020 presidential election voting

We will use multiple linear regression to build  models that investigate whether any of these variables can be useful predictors of body cameras being turned on during fatal shootings in the United States.

Because most states' body camera laws were enacted at the start of 2021, we will only look at data from 2021 and 2022. This reduces the number of cases in our original data set to 1,763.

First let us take a look at the new data set with its new variables (added in Part 3):

```{r}
#FINAL PROJECT WORK: ADDING NEW COLUMNS OF DATA   
FD <- subset(By_State, (year== 2022| year==2021) )
FD$regions = factor(FD$regions)
head(FD)
```

Number of observations:

```{r}
nrow(FD)
```

#### Figure 4: Body Camera Laws

The following figure depicts the relationship between body camera laws and stbcp. 

```{r}
#stbcp vs le_per_100k
ggplot(data=FD,aes(x = stbcp,y=bclaw)) + geom_point(aes(color=le_per_100k,size=le_per_100k))+ labs(x = "stbcp", y="Laws governing body camera usage")
```

#### Figure 5: Law Enforcement Officers Per 100,000 Citizens

The following figure depicts the relationship between law enforcement officers per 100K citizens and stbcp. 

```{r}
#stbcp vs le_per_100k
ggplot(data=FD,aes(x = stbcp,y=le_per_100k)) + geom_point(aes(color=le_per_100k,size=le_per_100k))+ labs(x = "stbcp", y="Law enforcment officers per 100,000 citizens")
```

#### Figure 6: 2020 Election Voting Margin

This figure depicts the relationship between the 2020 US presidential election margin and stbcp.

```{r}
#stbcp vs marg2020
ggplot(data=FD,aes(x = stbcp,y=marg2020)) + geom_point(aes(color=marg2020,size =marg2020))+ labs(x = "stbcp", y="Percent margin voting swing")
```

#### Figure 7: Spending on Policing Per Capita

This figure depicts the relationship between the state spending on policing per capita and stbcp.

```{r}
#spendpc vs regions
ggplot(data=FD,aes(x = regions,y=spendpc)) + geom_point(aes(color=regions,size =regions))+ labs(x = "stbcp", y="State spending on policing per capita")

```

#### Figure 8: Law Enforcement Officers Per 100,000 Citizens

This figure shows law enforcement officers per 100K citizens, grouped by region.

```{r}
#regions vs le_per_100k
ggplot(data=FD,aes(x = regions,y=le_per_100k)) + geom_point(aes(color=regions,size=regions))+ labs(x = "Region", y="Law enforcment officers per 100,000 citizens")
```

#### Figure 9: 2020 Election Voting Margin

The following plot compares the following variables: stbcp, 2020 election margin, and region. 

```{r}
fig <- plot_ly(FD, x = FD$regions, y = FD$stbcp, z = FD$marg2020,color =  FD$regions, colors = c('purple', 'yellow'))
fig <- fig %>% add_markers()
fig <- fig %>% layout(scene = list(xaxis = list(title = 'Regions'),
                     yaxis = list(title = 'Body Camera Proportion'),
                     zaxis = list(title = '2020 Election Margin')))
fig
```

#### Figure 10: Law Enforcement Officers Per 100,000 Citizens

Finally, here is a plot showing law enforcement officers per 100K citizens, police spending, and region.

```{r}
fig <- plot_ly(FD, x = FD$regions, y = FD$spendpc, z = FD$le_per_100k,color =  FD$regions, colors = c('blue', 'green'))
fig <- fig %>% add_markers()
fig <- fig %>% layout(scene = list(xaxis = list(title = 'Region'),
                     yaxis = list(title = 'Law Enforcement Spending'),
                     zaxis = list(title = 'Police per 100,000 citizens')))
fig
```

Now that we are familiar with the data, we can start to model with our new state-wide data.

#### Model 2: A simple MLRG model that uses all the new variables along with the region variable

```{r}
lm1 <- lm(stbcp~ (marg2020 + bclaw + regions + le_per_100k+ spendpc), data=FD)
summary(lm1)
car::vif(lm1 )
```
```{r}
# Grab residuals
res <- residuals(lm1)

# Convert to DataFrame for gglpot
res <- as.data.frame(res)

head(res)

```
```{r}
# Histogram of residuals
ggplot(res,aes(res)) +  geom_histogram(fill='blue',alpha=0.5)
```
```{r}

plot(lm1)
```

The VIF values for model 1 are all within acceptable range.

With an R^2 of 0.183, this model is not very good at predicting statewide body camera usage. We can see that the region variable is not helpful so we will remove it.

#### Model 3: Uses only the most helpful predictors from the previous model

```{r}
lm2 <- lm(stbcp~ (marg2020+bclaw+spendpc+ le_per_100k), data=FD)
summary(lm2)
car::vif(lm2)
```
```{r}
# Grab residuals
res <- residuals(lm2)

# Convert to DataFrame for gglpot
res <- as.data.frame(res)

head(res)
```
```{r}
# Histogram of residuals
ggplot(res,aes(res)) +  geom_histogram(fill='blue',alpha=0.5)
```
```{r}
plot(lm2)
```

The VIF values for model 2 are all within acceptable range.

With an R^2 of 0.152, this model is even worse at predicting statewide body camera usage.

#### Model 4: Interaction of law enforcement spending per capita on quantity of officers per capita

```{r}
#Redo of lm1 but with an interaction
lm3 <- lm(stbcp~ (marg2020 + bclaw + regions + le_per_100k+ spendpc + I(spendpc*le_per_100k)), data=FD)
summary(lm3)
```
```{r}
# Grab residuals
res <- residuals(lm3)

# Convert to DataFrame for gglpot
res <- as.data.frame(res)

head(res)
```
```{r}
# Histogram of residuals
ggplot(res,aes(res)) +  geom_histogram(fill='blue',alpha=0.5)
```
```{r}
plot(lm3)
```

We are ignoring the VIF test for multicolinearity because we are using an interaction predictor.

With an R^2 of 0.276, this model is not good, much better than the others at predicting statewide body camera usage.

#### Model 5: Testing lm3

Since lm3 is our best model (per our R^2), lets try to predict a few made up US states:

Please notice Eleum and Faraam are identical except their body camera laws. This is the same for GW and HW.

```{r}
Atlantis <- data.frame(bclaw = c(1), marg2020=c(-10), regions=c('SW'), le_per_100k= c(500), spendpc=c(720))
Burk <- data.frame(bclaw = c(0), marg2020=c(-33), regions=c('MW'), le_per_100k= c(249), spendpc=c(398))
Caelid <-data.frame(bclaw = c(1), marg2020=c(29), regions=c('NE'), le_per_100k= c(610), spendpc=c(880))
Dollywood <- data.frame(bclaw = c(0), marg2020=c(5), regions=c('SE'), le_per_100k= c(200), spendpc=c(203))

Eleum <- data.frame(bclaw = c(0), marg2020=c(-20), regions=c('NW'), le_per_100k= c(440), spendpc=c(545))
Faraam <- data.frame(bclaw = c(1), marg2020=c(-20), regions=c('NW'), le_per_100k= c(440), spendpc=c(545))

GW <- data.frame(bclaw = c(0), marg2020=c(20), regions=c('SE'), le_per_100k= c(540), spendpc=c(645))
HW <- data.frame(bclaw = c(1), marg2020=c(20), regions=c('SE'), le_per_100k= c(540), spendpc=c(645))
```

Now let's plug these new "states" into our model:

```{r}
predict(lm3, newdata = Atlantis, interval = 'confidence')
predict(lm3, newdata= Burk ,interval = 'confidence')
predict(lm3, newdata = Caelid, interval = 'confidence')
predict(lm3, newdata = Dollywood, interval = 'confidence')
predict(lm3, newdata = Eleum, interval = 'confidence')
predict(lm3, newdata = Faraam, interval = 'confidence')
predict(lm3, newdata = GW, interval = 'confidence')
predict(lm3, newdata = HW, interval = 'confidence')
```

We can see the difference of fit on states Eleum and Faraam, as well as GW and HW.

### Part 6: Conclusion

Studying the use of body cameras in police work is an important topic of study for data-driven policy research in the United States. While we hoped to be able to associate this correlation between the U.S. state of observation and whether the body camera was on or off during the shooting to state policy on body cameras to some variable, we were unable to find a strong correlation. Although lm3 was our best model, it is still not a great predictor of statewide body camera usage, which can lead us to the following conclusions:

1. Both regional and state groupings demonstrated quantifiable differences in the proportion of body cameras turned on or off during fatal police shootings.

2. The number of law enforcement officers per capita does not influence whether a body camera is turned on or off.

3. State spending on policing per capita does not influence whether a body camera is turned on or off.

4. The political affiliation of a states does not influence whether a body camera is turned on or off.

5. Body camera mandate laws, present in Maryland, New Jersey, New Mexico, and South Carolina, slightly influence whether a body camera is turned on or off. 

Considering body camera laws are relatively nascent, it will be an interesting topic of study to evaluate changes in stbcp as more states adopt such laws. This research project has shown that they may be the best chance states have of increasing camera usage during active police duty.
